Master's Project Journal

### 2019-04-16 thesis/NetworkImplemenation.tex
50.1 pages

\subsubsection{Frequency Divider}
The frequency divider is implemented using ``Divider 2'' from Chapter \ref{chap:3}, as only divisions ratios of 2, 4 and 8 were selected for use in the testing process. For ease of testing, the divider had outputs representing each of these division ratios simultaneously, which could be multiplexed between at runtime if so desired. A fourth output passed the input through unmodified, allowing for the divider to be removed without regeneration of the \ac{FPGA} configuration.
\begin{figure}[h]%
	\centering
	\includegraphics[width=0.6\textwidth]{../divider2}
	\caption[Frequency divider \ac{RTL} diagram]{Frequency divider \ac{RTL} diagram.}
	\label{fig:divs_iompl}
\end{figure}

@@

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{../ro_new}
	\caption[\acl{RO} RTL diagram]{\acl{RO} RTL diagram.}
	\label{fig:ro_impl}
\end{figure}
Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented. %TODO impl & discuss

A control code width of 5 bits was chosen as with two inverters removed per control code, the tunable range would consist of $20\%$ of the total inverter count. To achieve this range with a centre of $317$ inverters, the maximum number of inverters in the chain was required to be $317+2\times 2^{5-1} = 349$, dropped to a minimum at $285$. The corresponding minimum and maximum frequencies then are:
\begin{align}
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\times 0.5t_{step}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.136~\si{\nano\second}} = 4.354~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.136~\si{\nano\second}} = 5.332~\si{\mega\hertz}
\end{align}

@@

When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delievered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$. For the purposes of describing a typical tuning range, the average propagation delay of an \ac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used. To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.

@@

\subsection{\acs{ADPLL} Design 3}
\subsubsection{\acl{DCO}}
\subsubsection{Phase Detector}

>>>>

\subsubsection{Frequency Divider}
The frequency divider is implemented using ``Divider 2'' from Chapter \ref{chap:3}, as only divisions ratios of 2, 4 and 8 were selected for use in the testing process. For ease of testing, the divider had outputs representing each of these division ratios simultaneously, which could be multiplexed between at runtime if so desired. A fourth output passed the input through unmodified, allowing for the divider to be removed without regeneration of the \ac{FPGA} configuration.
\begin{figure}[h]%
	\centering
	\includegraphics[width=0.6\textwidth]{../divider2}
	\caption[Frequency divider \ac{RTL} diagram]{Frequency divider \ac{RTL} diagram.}
	\label{fig:divs_impl}
\end{figure}

@@

\subsubsection{\acs{ADPLL} 1 Design Summary}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{../ro_new}
	\caption[\acl{RO} RTL diagram]{\acl{RO} RTL diagram.}
	\label{fig:ro_impl}
\end{figure}
Figure \ref{fig:ro_impl} contains an \ac{RTL} diagram of the \ac{RO} as implemented. The inverter chain is specified with a maximum length of $n$ and a maximum number of removable inverters of $k$. In order to preserve the unstable circuit, by maintaining an odd number of inverters, only multiples of two inverters may be removed at a time. The already introduced \ac{DCO} and \ac{PFD} designs all followed the same convention regarding phase error and the corresponding impact on the control code: In the \ac{PFD}, if the generated signal was lagging the reference, meaning it should ``go faster'' in order to ``catch up'', the error given a positive sign. Similarly in the \ac{FPGA} clocked \ac{DCO}, a positive change in the magnitude of the control code resulted in a higher frequency. In the interests of consistency and extensibility the same behaviour was implemented for this \ac{DCO} also, with the number of inverters in the chain given as:
\begin{equation}
\text{num\_inverters} = inv_{max} - 2\times\text{control\_code},~\text{control\_code}<\frac{k}{2}
\end{equation}
$k$ is determined by the bit width of the control code which, as in the case of the clocked \ac{DCO}, is also an unsigned number. An inverter chain will oscillate freely and cannot be disabled. To this end the first inverter in the chain was replaced by a \ac{NAND} gate to grant control over its operation.

The bias point functions differently in this oscillator, as unlike the \ac{FPGA} clocked design it is not required to set the centre of the tuning range, as in the \ac{RO} this lies at $n-\frac{k}{2}$. Instead the bias point servers as a conversion between the signed \ac{LF} output, which is in two's complement form, and the unsigned requirement of the control code. This is achieved by adding on the midpoint of the unsigned range and storing the result in an unsigned value. Using the case where the width of the control code is three: The range of the signed three bit integer is from $-4$ (100) to $3$ (011) which when directly converted to unsigned values become $4$ and $3$, thus destroying the relationship between lead and lag. Worse still the signed value of $-1$ (111) becomes $7$. Adding on the mid range value solves this problem, with a signed value of $-4$ now mapping to $0$ and $3$ to $7$. Comparing this behaviour to response of the \ac{RO} to the control code, it can be trivially seen that the minimum possible period will still correspond to the largest control code prior to conversion and vice versa for the maximum period.

A control code width of 5 bits was chosen, as with two inverters removed per control code, the tunable range would consist of $20\%$ of the total inverter count. To achieve this range with a centre of $317$ inverters, the maximum number of inverters in the chain was required to be $317+2\times 2^{5-1} = 349$, dropped to a minimum at $285$. The corresponding minimum and maximum frequencies then are:
\begin{align}
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}} = \frac{1}{\text{num. inverters}\times 0.5t_{step}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{349\times 0.5\times1.26~\si{\nano\second}} = 4.548~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{285\times 0.5\times1.26~\si{\nano\second}} = 5.569~\si{\mega\hertz}
\end{align}

@@

When, later in the project, additional \ac{RO}s were implemented to form networks, 373 inverters proved a suitable starting point which consistently delivered \ac{DCO}s with a centre frequency within $200~\si{\kilo\hertz}$ of $5~\si{\mega\hertz}$. For the purposes of describing a typical tuning range, the average propagation delay of an \ac{RO} consisting of 373 inverters with a centre frequency of $5.06~\si{\mega\hertz}$ will be used. To avoid the aforementioned issues with constant control codes, the fixed code was achieved by implementing an \ac{ADPLL} and setting the \ac{PI} filter gains to zero at runtime before performing a reset.

Using this updated maximum number of inverters the frequency range can be recalculated. Firstly, the new minimum and maximum numbers of inverters are $309$ and $373$ respectively, with a mid point at $341$. The inverter delay and period step are then recomputed based on the centre frequency of $5~\si{\mega\hertz}$ as:
\begin{align}%TODO 2 x 2 x midpoint justification - in prev. chapter?
t_{step} = \frac{T_{osc}}{0.5\times\text{inverter midpoint}} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second}
\text{propagation delay} &= \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second} \\
\end{align}
The frequency range is then:
\begin{align} %TODO inconsisentl calculation approach with previous version
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{\text{num. inverters}\times 2\times\text{propagation delay}}
f_{min} &= \frac{1}{T_{max}} = \frac{1}{373\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{309\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
\end{align}

@@

\subsubsection{\acs{ADPLL} 2 Design Summary}

\subsection{\acs{ADPLL} Design 3}
\subsubsection{\acl{DCO}}
\subsubsection{Phase Detector}
\subsubsection{\acs{ADPLL} 1 Design Summary}
